---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Take care of some housekeeping 
```{r}
# following this site
rm(list=ls())
dir.create(file.path("./", "Generated.Results"), showWarnings = FALSE)
```
Install some packages:
```{r}
library("phyloseq")
library("vegan")
library("ggplot2")
#library("tidyr")
library("dplyr")
#library("automap")
#library("mltools")
```
The working directory is set within this RMD package, 
But we're still going to need some data to work with
For now this is stored on a local repository 
In the future link this to a figshare DOI
```{r}
#set WD
setwd("/Users/fungi/Google Drive/Projects/Vanuatu/Vanuatu Molecular/New_Analysis/R")

# load data
# a transformed phyloseq object 
vst_physeq <- readRDS(file = "vst_physeq_geog.rds")
```

Remove OTUs that do not show appear more than 0 times in more than half the samples
```{r}
sbst <- subset_samples(vst_physeq, Sample_Type == c("Understory","Ground"))
    count_tab <-  otu_table(sbst) #this is an OTU table with abundences for the OTUs subsetted above 
    count_tab[count_tab <= 0.0] <- 0.0 # since this is a VST transeformed dataset, we'll need to set negative values to zero

    count_tab <- count_tab[rowSums(count_tab)>1,]
    
     otu_table(sbst) <- count_tab # and reinsert into the phylloseq object 
 sample_sums(sbst)[1:10] # double check to make sure tha worked 
```

```{r}
GP.ord <- ordinate(sbst, "NMDS", "bray")
```
Only if you're interested in plotting by transect also 
```{r}
class(sample_data(sbst)$Transect)
sample_data(sbst)$Transect <- as(sample_data(sbst)$Transect, "character")
```

```{r}
p2 = plot_ordination(sbst, GP.ord,color="Sample_Type")#, shape = "Transect")
p2 + 
  #geom_polygon(aes(fill=Sample_Type)) + 
  geom_point(size=2) + 
  stat_ellipse() +
 theme_bw() 
 #labs(title = "NMDS Plot")
```
