---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(plyr)
```

Check to see if there is a directory that can hold all the generated results. If not, make one
```{r}
dir.create(file.path("./", "Generated.Results"), showWarnings = FALSE)
```



Set the minimum number of occurences an OTU must appear in a transect

This is not so important within this rmd code, but it was really important while setting up the project. 
See "./Users/fungi2/Google Drive/Projects/Vanuatu/Vanuatu Molecular/New_Analysis/R/Mapping.Single.Species.Figure3.lm-pres_occ.R""
```{r}
BBB=2
```

```{r}
#setwd("/Users/fungi2/Google Drive/Projects/Vanuatu/Vanuatu Molecular/New_Analysis/R")



# Load up the required CSV
BBB=2
gr.spec <- read.csv(file = paste("~/Google Drive/Projects/Vanuatu/Vanuatu Molecular/New_Analysis/R/Mapping.single.sp/Mantel.0.05/Patch.occupancy/special.grnd.",BBB,".occur.results.csv"), header = TRUE)
un.spec <- read.csv(file = paste("~/Google Drive/Projects/Vanuatu/Vanuatu Molecular/New_Analysis/R/Mapping.single.sp/Mantel.0.05/Patch.occupancy/special.undr.",BBB,".occur.results.csv"), header = TRUE)
ungr.gen <- read.csv(file = paste("~/Google Drive/Projects/Vanuatu/Vanuatu Molecular/New_Analysis/R/Mapping.single.sp/Mantel.0.05/Patch.occupancy/gen.grnd.undr.",BBB,".occur.results.csv"), header = TRUE)

par(mfrow=c(2,2))
scatter.smooth(x=gr.spec$Ground.PA, y=gr.spec$Ground.abund/gr.spec$Ground.PA, 
               main="Soil Specialist", 
               xlab="Transect occupancy", ylab="Abundance/Occupancy",
               lpars = list(col = "red", lwd = 3, lty = 3), pch = 3) 
linearMod <- lm(Ground.abund/gr.spec$Ground.PA ~ Ground.PA, data=gr.spec)  # build linear regression model on full data
summary(linearMod)

scatter.smooth(x=un.spec$Under.PA, y=un.spec$Under.abund/un.spec$Under.PA, 
               main="Phyllosphere Specialist", 
               xlab="Transect occupancy", ylab="Abundance/Occupancy",
               lpars = list(col = "red", lwd = 3, lty = 3), pch = 3) 
linearMod <- lm(Under.abund ~ Under.PA, data=un.spec)  # build linear regression model on full data
summary(linearMod)


scatter.smooth(x=ungr.gen$Ground.PA, y=ungr.gen$Ground.abund/ungr.gen$Ground.PA, 
               main="Soil Generalists", 
               xlab="Transect occupancy", ylab="Abundance/Occupancy",
               lpars = list(col = "red", lwd = 3, lty = 3), pch = 3) 
linearMod <- lm(Ground.abund ~ Ground.PA, data=ungr.gen)  # build linear regression model on full data
summary(linearMod)

scatter.smooth(x=ungr.gen$Under.PA, y=ungr.gen$Under.abund/ungr.gen$Under.PA,
               main="Phyllosphere Generalists", 
               xlab="Transect occupancy", ylab="Abundance/Occupancy",
               lpars = list(col = "red", lwd = 3, lty = 3), pch = 3) 
linearMod <- lm(Ground.abund ~ Ground.PA, data=ungr.gen)  # build linear regression model on full data
summary(linearMod)
dev.off()


####################################################################
gr.spec2 <- gr.spec[,1:3]
gr.spec2$ID <- "Soil.Specialist"
gr.spec2$Habitat <- "Soil"
gr.spec2$Lifestyle <- "Specialist"
names(gr.spec2) <- c("OTU","Abundance","Occupancy","ID","Habitat","lifestyle")

un.spec2 <- un.spec[,c(1,4:5)]
un.spec2$ID <- "Phyllo.Specialist"
un.spec2$Habitat <- "Phyllosphere"
un.spec2$Lifestyle <- "Specialist"
names(un.spec2) <- c("OTU","Abundance","Occupancy","ID","Habitat","lifestyle")

#ungr.gen 
gr.gen <- ungr.gen[,1:3]
gr.gen$ID <- "Soil.Generalist"
gr.gen$Habitat <- "Soil"
gr.gen$Lifestyle <- "Generalist"
names(gr.gen) <-  c("OTU","Abundance","Occupancy","ID","Habitat","lifestyle")

un.gen <- ungr.gen[,c(1,4:5)]
un.gen$ID <- "Phyllo.Generalist"
un.gen$Habitat <- "Phyllosphere"
un.gen$Lifestyle <- "Generalist"
names(un.gen) <-  c("OTU","Abundance","Occupancy","ID","Habitat","lifestyle")


full.table <- rbind(gr.spec2,un.spec2,gr.gen,un.gen)
full.table$Avg.Abund <- full.table$Abundance/full.table$Occupancy


#Set colour palette
ground.und.colours <- c( "#669900","#996600")
spec.gen.colours <-  c("#404080", "#69b3a2")

names(full.table)
```
Facet by specialization
```{r}
ggplot(data=full.table, aes(x=Occupancy, y=Avg.Abund, col=Habitat))+
  geom_point(alpha = 4/10, size = 0.7)+
  scale_colour_manual(values=ground.und.colours)+
  stat_smooth(method='loess')+
  facet_grid(.~lifestyle)+
  #scale_color_viridis() +
  #labs(title = "Interspecific Occupancy–abundance relationship") +
  xlab(paste("Transect occupancy"))+ # \n (number of occurences / transect)")) +
  ylab(paste("Local abundance"))+ # \n (OTU sequence reads/number of occurences)")) +
    theme(panel.background = element_rect(fill = "white", colour = "black"),
        strip.background = element_rect(fill = "white", colour = "black"),
        #legend.key = element_blank(),
        legend.justification = c(1.1, 0),# x-axis 
        legend.position = c(1, 0.75),# y-axis
        legend.title = element_blank(),
        strip.text.x = element_text(size=12, face="bold"))

```

And save it 
```{r}
ggsave( plot = last_plot(), filename = paste("./Generated.Results/Occupancy_abundance.MinTransectOccur",BBB,".pdf"), device = NULL, path = NULL,
        scale = 1, width = 180, height = 100, units = "mm", 
        dpi = 300, limitsize = TRUE)
```



Facet by habitat
```{r}
ggplot(data=full.table, aes(x=Occupancy, y=Avg.Abund, col=lifestyle))+
  geom_point(alpha=0.3, size = 0.7)+
  scale_colour_manual(values=spec.gen.colours)+
  stat_smooth(method='loess')+
  facet_grid(.~Habitat)+
  #scale_color_viridis() +
  #labs(title = "Interspecific Occupancy–abundance relationship") +
  xlab(paste("Transect occupancy"))+ # \n (number of occurences / transect)")) +
  ylab(paste("Local abundance"))+ # \n (OTU sequence reads/number of occurences)")) +
    theme(panel.background = element_rect(fill = "white", colour = "black"),
        strip.background = element_rect(fill = "white", colour = "black"),
        #legend.key = element_blank(),
        legend.justification = c(1.1, 0),# x-axis 
        legend.position = c(1, 0.75),# y-axis
        legend.title = element_blank(),
        strip.text.x = element_text(size=15, face="bold"))
```
And save it 
```{r}
ggsave( plot = last_plot(), filename = paste("./Generated.Results/Fig5.OA.pdf"), device = NULL, path = NULL,
        scale = 1, width = 180, height = 100, units = "mm", 
        dpi = 300, limitsize = TRUE)
```






#Non-parametric tests
Do not report means and standard deviations for non-parametric tests. Report the median
and range in the text or in a table


```{r}

mu <- ddply(full.table, "ID", summarise,# grp.mean=mean(Occupancy))
 median = round(median(Occupancy), 2),
  IQR = round(IQR(Occupancy), 2),
 N = round(length(Occupancy), 2),
  mean = round(mean(Occupancy), 2),
 sd = round(sd(Occupancy), 2))


print(mu)
```
Test for differences
```{r}

pairwise.wilcox.test(full.table$Occupancy,full.table$ID,p.adjust.method = "BH")
#wilcox.test(Occupancy ~ lifestyle, data = full.table, exact = FALSE)
```


#Area occupancy 

We're going to make a histogram of Occupancy x OTU count 
Separated by habitat and life history 


```{r}
#colo <-c("#F0E442","#009E73") #Printer friendly / colour blind frinedly colours 
colo <- c("#404080", "#69b3a2","#404080", "#69b3a2") #These are also nice 


ggplot(data = full.table, aes(x=Occupancy, fill=lifestyle)) +
    geom_histogram(binwidth = 3, alpha=0.3, position = 'identity')+
  scale_fill_manual(values=colo)+
  #geom_vline(data=mu, aes(xintercept=mean, color=lifestyle),
   #          linetype="dotted", color=colo, size=1)+
   scale_y_continuous(trans = "log10") +
   facet_wrap(~Habitat, scales = 'free_x') +
   xlab(paste("Transect occupancy"))+ # \n (number of occurences / transect)")) +
   ylab(paste("Observed species"))+ # \n (OTU sequence reads/number of occurences)")) +
   theme(panel.background = element_rect(fill = "white", colour = "black"),
        strip.background = element_rect(fill = "white", colour = "black"),
        #legend.key = element_blank(),
        legend.justification = c(1.1, 0),# x-axis 
        legend.position = c(1, 0.75),# y-axis
        legend.title = element_blank(),
        strip.text.x = element_text(size=12, face="bold"))

```


And save it 
```{r}
ggsave( plot = last_plot(), filename = paste("./Generated.Results/Area_occurence.pdf"), device = NULL, path = NULL,
        scale = 1, width = 180, height = 100, units = "mm", 
        dpi = 300, limitsize = TRUE)
```





